Symmetric Weighted First-Order Model Counting and Factorial-Like Functions

### 2021-10-06 talk/talk.tex
Fix a bug with domain size reduction. Partial injection counting now works!
Injection counting (but exponential...)

\begin{frame}{Counting (Unweighted) Functions: Currently Unliftable (1)}
  \cite{stanley2011enumerative}
  \begin{block}{Functions $M \to N$ (let $|M| = m$ and $|N| = n$)}
    Theory:
    \begin{align*}
      \forall x \in M. \forall y, z \in N. &P(x, y) \land P(x, z) \Rightarrow y=z \\
      \forall x \in M. \exists y \in N. &P(x, y)
    \end{align*}
    Answer: $n^m$.
  \end{block}
  \begin{block}{Injections}
    Theory:
    \begin{align*}
      \forall x \in M. \forall y, z \in N. &P(x, y) \land P(x, z) \Rightarrow y=z \\
      \forall x \in M. \exists y \in N. &P(x, y) \\
      \alert{\forall w, x \in M. \forall y \in N.} &\alert{P(w, y) \land P(x, y) \Rightarrow w=x}
    \end{align*}
    Answer: $(n)_m = n \cdot (n-1)\cdots(n-m+1)$.
  \end{block}
\end{frame}

@@

\begin{frame}{Counting (Unweighted) Functions: Currently Unliftable (3)}
  These ones could contain errors.
  \begin{block}{Partial functions}
    Theory: $\forall x \in M. \forall y, z \in N. P(x, y) \land P(x, z)
    \Rightarrow y=z$

    Answer: $(n+1)^m$.
  \end{block}
  \begin{block}{Partial injections}
    Theory:
    \begin{align*}
      \forall x \in M. \forall y, z \in N. &P(x, y) \land P(x, z) \Rightarrow y=z \\
      \forall w, x \in M. \forall y \in N. &P(w, y) \land P(x, y) \Rightarrow w=x
    \end{align*}
    Answer: $\sum_{k=0}^{\min\{m, n\}} k!\binom{m}{k}\binom{n}{k}$.
  \end{block}
\end{frame}

>>>>

\begin{frame}{Counting (Unweighted) Functions: Currently Unliftable (1)}
  \cite{stanley2011enumerative}
  \begin{block}{Functions $M \to N$ (let $|M| = m$ and $|N| = n$)}
    Theory:
    \begin{align*}
      \forall x \in M. \forall y, z \in N. &P(x, y) \land P(x, z) \Rightarrow y=z \\
      \forall x \in M. \exists y \in N. &P(x, y)
    \end{align*}
    Answer: $n^m$.
  \end{block}
  \begin{block}{Injections}
    Theory:
    \begin{align*}
      \forall x \in M. \forall y, z \in N. &P(x, y) \land P(x, z) \Rightarrow y=z \\
      \forall x \in M. \exists y \in N. &P(x, y) \\
      \alert{\forall w, x \in M. \forall y \in N.} &\alert{P(w, y) \land P(x, y) \Rightarrow w=x}
    \end{align*}
    Answer: $n^{\underline{m}} = n \cdot (n-1)\cdots(n-m+1)$ if $m \le n$ and $0$ otherwise (for positive $m$ and $n$).
  \end{block}
\end{frame}

@@

\begin{frame}{Counting (Unweighted) Functions: Currently Unliftable (3)}
  \begin{block}{Partial functions}
    Theory: $\forall x \in M. \forall y, z \in N. P(x, y) \land P(x, z)
    \Rightarrow y=z$

    Answer: $(n+1)^m$.
  \end{block}
  \begin{block}{Partial injections}
    Theory:
    \begin{align*}
      \forall x \in M. \forall y, z \in N. &P(x, y) \land P(x, z) \Rightarrow y=z \\
      \forall w, x \in M. \forall y \in N. &P(w, y) \land P(x, y) \Rightarrow w=x
    \end{align*}
    My answer: $\sum_{k=0}^{\min\{m, n\}} k!\binom{m}{k}\binom{n}{k}$.

    \pause
    \alert{Answer found by Forclift:}
    \[
    f(m, n) =
    \begin{cases}
      1 & \text{if } m = 0 \\
      \sum_{k=0}^n \binom{n}{k} [k < 2] f(m-1, k) & \text{otherwise.}
    \end{cases}
    \]
    (exponential...)
  \end{block}
\end{frame}
